Taking Dyads Seriously

### 2021-01-06 paper/1_intro.tex
rev1 notes

The aim of this work is to address how to estimate the effect of exogenous covariates in a GLM context when there are network dependencies between observations in dyadic data. \citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).  However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid). Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases. In practice, such strategies remain insufficient because they do not directly model underlying structural features in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data.''} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).} The aims of our study are twofold. First, we provide a framework that accounts for dependencies often present in dyadic data.\footnote{This is important to consider since the inferential problems caused by non-iid observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.} Second, the framework we discuss here enables applied scholars to understand the structure underlying dyadic data rather than simply controlling for them.

>>>>

The aim of this work is to address how to estimate regression coefficients in a GLM context when there are network dependencies between observations in dyadic data. \citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).  However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid). Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases. In practice, such strategies remain insufficient because they do not directly model underlying structural features in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data.''} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).} The aims of our study are twofold. First, we provide a framework that accounts for dependencies often present in dyadic data.\footnote{This is important to consider since the inferential problems caused by non-iid observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.} Second, the framework we discuss here enables applied scholars to understand the structure underlying dyadic data rather than simply controlling for them.
